Indirect elicitation as a necessary condition for consistent surrogate losses

### 2020-05-28 neurips-20/cvx-flats.tex
Added a finite example where our bound strictly beats FSD

	As $L'$ is calibrated with respect to $\ell'$ and indirectly elicits $\gamma' := \prop{\ell'}$, then by Theorem~\ref{thm:cvx-flats}, we know there exists a flat $F'$ of codimension $d$ relative to $\reals^{\Y'}$ so that $\dim(F') = \|p\|_0 - d - 1$.
	Moreover, as $d$ is the maximal codimension of such a flat, we show in Lemma~\ref{lem:feas-sub-is-a-flat} that $\S_{\gamma_r}(p)$ is one such flat, and we have $\dim(\S_{\gamma_r}(p)) \geq \dim(F') = \|p\|_0 - d- 1$.
	Lemma~\ref{lem:p-boundary-fsd} then states $\dim(\S_{\gamma'_r}(p)) = \dim(\S_{\gamma_r}(p))$, and from there we observe the result.
%	First, for intuition, consider $p \in \relint{\simplex}$. 
%	Theorem~\ref{thm:cvx-flats} implies that there exists a flat $F'$ of affine dimension at least $n-d-1$.
%	Lemma~\ref{lem:feas-sub-is-a-flat} then says that $\dim(\S_{\gamma_r}(p)) \geq \dim(F') \geq n-d-1 \implies d \geq n - \dim(\S_{\gamma_r}(p)) - 1$.
%
%	
%	If we can show that $L' := L|_{\Y'}$ is calibrated with respect to $\ell' := \R \times \Y' \to \reals$ with $\ell(r,y) = \ell'(r,y)$ for all $r \in \R$ and $y \in \Y'$ and indirectly elicits $\gamma' := \prop{\ell'}$, then we observe the existence of a flat $F^*$ of codimension $d$ (in $\reals^{\Y'}$) so that $\dim(F^*) = \|p\|_0 - d - 1$ by Theorem~\ref{thm:cvx-flats}.
%	Then we can use Lemma~\ref{lem:p-boundary-fsd} to observe $\dim(\S_{\gamma'_r}(p)) = \dim(\S_{\gamma_r}(p))$, and thus the result holds.
%	
%	Now to see that $L'$ is calibrated with respect to $\ell'$, consider that $\gamma(p) = \gamma'(p)$ for all $p \in \Delta^{\Y'}$,
%	\begin{align*}
%		\forall p \in \Delta_{\Y'}:  \inf_{u \in \reals^d : \psi(u) \not \in \gamma'(p)} L'(u;p) = \inf_{u \in \reals^d : \psi(u) \not \in \gamma(p)} L(u;p) > \inf_{u \in \reals^d} L(u;p) = \inf_{u \in \reals^d} L'(u;p)~.~
%	\end{align*}
%	Therefore, we have $L'$ calibrated with respect to $\ell'$.
%
%	Moreover, we want to show $L'$ indirectly elicits $\gamma'$ so we can apply Theorem~\ref{thm:cvx-flats}.
%	First, observe that $L'$ elicits $\Gamma' : p \mapsto \Gamma(p)$ for all $p \in \Delta_{\Y'}$ 
%	For all $p \in \simplex$, we have $u \in \Gamma(p) \implies \psi(u) \in \gamma(p)$.
%	As $\Delta_{\Y'} \subseteq \simplex$, this is in particular true for all $p \in \Delta_{\Y'}$.
%	Therefore, for all $p \in \Delta_{\Y'}$, we have $\Gamma'_u = \Gamma_u \cap \Delta_{\Y'} \subseteq \gamma_{\psi(u)} \cap \Delta_{\Y'} = \gamma'_{\psi(u)}$, and thus $L'$ indirectly elicits $\gamma'$.
\end{proof}

@@

Theorem~\ref{thm:cvx-flats} also addresses one major open question from~\cite{frongillo2015elicitation,frongillo2018elicitation}\jessiet{which citation?}: \emph{what are lower bounds on convex elicitation complexity?}
Since proposition~\ref{prop:consistent-implies-indir-elic-prop} allows us to use convex elicitation complexity as a tool to understand efficiency of consistent convex surrogates for a given property, which is more often what is given in a continuous prediction setting.

For example, when one wants to learn an $\alpha$-quantile, we start with the property rather than a loss.
In the literature, when one wants to learn a quantile, pinball loss $L(r,y) = (r-y)(\mathbbm{1}_{r \geq y} - \alpha)$ typically appears without explanation or justification.
Elicitation allows us to understand why pinball loss is consistent for learning a quantile as the pinball loss elicits the $\alpha$-quantile.

@@

\begin{definition}
  A property $\Gamma:\simplex\to\R$ is \emph{$d$-identifiable} if its level sets are flats of co-dimension at most $d$ intersected with $\simplex$.
  \raft{Update to be consistent with our definitions / notation / conventions.  I'd vote for using co-dimension since it makes so much of the analysis easier...  Could make more formal too.}
  We write $\iden(\Gamma) = \min\{d \in \mathbb{N} : \Gamma\text{ is $d$-identifiable}\}$.
\end{definition}
\jessiet{Flats or unions of flats?}

\raf{A modification of the argument of \citet[Corollary 7]{frongillo2018elicitation}}

\begin{theorem}
  \label{thm:bayes-risk-lower-bound}
  Let $L$ elicit some identifiable $\Gamma:\simplex\to\reals^d$ with $\iden(\Gamma)=d$.
  If $\lbar$ is non-constant on every non-singleton level set of $\Gamma$, then $\eliccvx(\lbar) \geq \min(d+1,n-1)$.
\end{theorem}
\begin{proof}
  Suppose for a contradiction that we have some convex-elicitable property $\hat\Gamma:\simplex\to\reals^d$ and link $\psi : \reals^d \to \reals$ such that $\lbar = \psi \circ \hat\Gamma$.
  The condition $\iden(\Gamma)=d$ implies that $\Gamma$ is $d$-identifiable but not $(d-1)$-identifiable, and thus there must be some level set $\Gamma_r$ which is a flat of dimension $n-d-1$.
  \raft{What I mean is you can't describe $\Gamma_r$ using a flat of another dimension; revisit once we settle terminology}

>>>>

	As $L'$ is calibrated with respect to $\ell'$ and indirectly elicits $\gamma' := \prop{\ell'}$, then by Theorem~\ref{thm:cvx-flats}, we know there exists a flat $F'$ of codimension $d$ relative to $\reals^{\Y'}$ so that $\dim(F') = \|p\|_0 - d - 1$.
	Moreover, as $d$ is the maximal codimension of such a flat, we show in Lemma~\ref{lem:feas-sub-is-a-flat} that $\S_{\gamma_r}(p)$ is one such flat, and we have $\dim(\S_{\gamma_r}(p)) \geq \dim(F') = \|p\|_0 - d- 1$.
	Lemma~\ref{lem:p-boundary-fsd} then states $\dim(\S_{\gamma'_r}(p)) = \dim(\S_{\gamma_r}(p))$, and from there we observe the result.
%	First, for intuition, consider $p \in \relint{\simplex}$. 
%	Theorem~\ref{thm:cvx-flats} implies that there exists a flat $F'$ of affine dimension at least $n-d-1$.
%	Lemma~\ref{lem:feas-sub-is-a-flat} then says that $\dim(\S_{\gamma_r}(p)) \geq \dim(F') \geq n-d-1 \implies d \geq n - \dim(\S_{\gamma_r}(p)) - 1$.
%
%	
%	If we can show that $L' := L|_{\Y'}$ is calibrated with respect to $\ell' := \R \times \Y' \to \reals$ with $\ell(r,y) = \ell'(r,y)$ for all $r \in \R$ and $y \in \Y'$ and indirectly elicits $\gamma' := \prop{\ell'}$, then we observe the existence of a flat $F^*$ of codimension $d$ (in $\reals^{\Y'}$) so that $\dim(F^*) = \|p\|_0 - d - 1$ by Theorem~\ref{thm:cvx-flats}.
%	Then we can use Lemma~\ref{lem:p-boundary-fsd} to observe $\dim(\S_{\gamma'_r}(p)) = \dim(\S_{\gamma_r}(p))$, and thus the result holds.
%	
%	Now to see that $L'$ is calibrated with respect to $\ell'$, consider that $\gamma(p) = \gamma'(p)$ for all $p \in \Delta^{\Y'}$,
%	\begin{align*}
%		\forall p \in \Delta_{\Y'}:  \inf_{u \in \reals^d : \psi(u) \not \in \gamma'(p)} L'(u;p) = \inf_{u \in \reals^d : \psi(u) \not \in \gamma(p)} L(u;p) > \inf_{u \in \reals^d} L(u;p) = \inf_{u \in \reals^d} L'(u;p)~.~
%	\end{align*}
%	Therefore, we have $L'$ calibrated with respect to $\ell'$.
%
%	Moreover, we want to show $L'$ indirectly elicits $\gamma'$ so we can apply Theorem~\ref{thm:cvx-flats}.
%	First, observe that $L'$ elicits $\Gamma' : p \mapsto \Gamma(p)$ for all $p \in \Delta_{\Y'}$ 
%	For all $p \in \simplex$, we have $u \in \Gamma(p) \implies \psi(u) \in \gamma(p)$.
%	As $\Delta_{\Y'} \subseteq \simplex$, this is in particular true for all $p \in \Delta_{\Y'}$.
%	Therefore, for all $p \in \Delta_{\Y'}$, we have $\Gamma'_u = \Gamma_u \cap \Delta_{\Y'} \subseteq \gamma_{\psi(u)} \cap \Delta_{\Y'} = \gamma'_{\psi(u)}$, and thus $L'$ indirectly elicits $\gamma'$.
\end{proof}
\subsection{Example}\label{subsec:examples-finite}
There are some cases where the bound provided by Theorem~\ref{thm:cvx-flats} is strictly tighter than the bound provided by feasible subspace dimension in Corollary~\ref{cor:fsd-bound}.
For example, we can consider the \emph{abstain property} over 3 outcomes where one wishes to report the most likely outcome $y$ if $Pr[Y=y|x] \geq 1/2$ and abstains from predicting ($\bot$) otherwise.
This is elicited by the target loss $\ell^{abs} := \mathbbm{1}\{r \not \in \{y, \bot\}\} + (1/2)\mathbbm{1}\{r = \bot\}$. 
For this example, we choose $p$ to be the uniform distribution.
In this case, the bound by feasible subspace dimension yields $\ccdim(\ell^) \geq 3 - 2 - 1 = 0$, as the feasible subspace dimension is $2$ since we are on the relative interior of the level set and simplex, as shown in Figure~\ref{fig:fsd-bound}.

However, consider any flat of codimension $1$ (a line in the two dimensions; a plane in three) containing $p$.
When intersected with the simplex, one can see that any line in the simplex through $p$ also leaves the cell $\gamma_\bot$, which contains $p$.
See Figure~\ref{fig:flats-bound} for intuition.
Therefore, we have no flat of codimension $1$ containing $p$ staying in $\gamma_\bot$, so we obtain $\ccdim(\ell^{abs}) \geq 3 - 0 - 1 = 2$.

\begin{figure}
\begin{minipage}{0.45\linewidth}
	\centering
	\includegraphics[width=\linewidth]{tikz/fsd-bound.pdf}
	\caption{Since the uniform distribution is on the relative interior of the simplex and the level set $\gamma_\bot$, its feasible subspace dimension is $2$.}
	\label{fig:fsd-bound}
\end{minipage}
\hfill
\begin{minipage}{0.45\linewidth}
	\centering
	\includegraphics[width=\linewidth]{tikz/flats-bound.pdf}
	\caption{Any flat of codimension $1$ through the uniform distribution also leaves the (gray) cell $\gamma_\bot$ in the simplex.}
	\label{fig:flats-bound}
\end{minipage}
\end{figure}

@@

Theorem~\ref{thm:consistent-implies-indir-elic} allows us to use convex elicitation complexity as a tool to understand efficiency of consistent convex surrogates for a given property, which is more often what is given in a continuous estimation setting.
For example, when one wants to learn an $\alpha$-quantile, we start with the property rather than a loss.
In the literature, pinball loss $L(r,y) = (r-y)(\mathbbm{1}_{r \geq y} - \alpha)$ typically appears without explanation or justification as to why it is relevant in relation to the $\alpha$-quantile.
Elicitation teaches us why pinball loss is consistent for learning a quantile as the pinball loss elicits the $\alpha$-quantile.

Theorem~\ref{thm:cvx-flats} also addresses one major open question from~\cite{frongillo2015elicitation,frongillo2018elicitation}\jessiet{which citation?}: \emph{what are lower bounds on convex elicitation complexity?}
While Theorem~\ref{thm:cvx-flats} is a statement about consistent surrogates, the heart of it is actually about indirect property elicitation.
Thus, the result also applies to extend previous results from~\cite{frongillo2015elicitation,frongillo2018elicitation} to provide new bounds on elicitation complexity for certain classes of properties in Theorem~\ref{thm:bayes-risk-lower-bound}.

@@

When extending the results of~\cite{frongillo2018elicitation}, we assume a property is identifiable, meaning its level sets are flats.
\begin{definition}
  A property $\Gamma:\simplex\to\R$ is \emph{$d$-identifiable} if its level sets are flats of co-dimension at most $d$ intersected with $\simplex$.
  \raft{Update to be consistent with our definitions / notation / conventions.  I'd vote for using co-dimension since it makes so much of the analysis easier...  Could make more formal too.}
  We write $\iden(\Gamma) = \min\{d \in \mathbb{N} : \Gamma\text{ is $d$-identifiable}\}$.
\end{definition}
\jessiet{Flats or unions of flats?}

\raf{A modification of the argument of \citet[Corollary 7]{frongillo2018elicitation}}
In the proof of Theorem~\ref{thm:consistent-implies-indir-elic} below, we modify the argument of~\cite[Corollary 7]{frongillo2018elicitation} to bound the elicitation complexity of the Bayes Risk of the loss.
\begin{theorem}
  \label{thm:bayes-risk-lower-bound}
  Let $L$ elicit some identifiable $\Gamma:\simplex\to\reals^d$ with $\iden(\Gamma)=d$.
  If $\lbar$ is non-constant on every non-singleton level set of $\Gamma$, then $\eliccvx(\lbar) \geq \min(d+1,n-1)$. 
\end{theorem}
\begin{proof}
  Suppose for a contradiction that we have some convex-elicitable property $\hat\Gamma:\simplex\to\reals^d$ and link $\psi : \reals^d \to \reals$ such that $\lbar = \psi \circ \hat\Gamma$.
  The condition $\iden(\Gamma)=d$ implies that $\Gamma$ is $d$-identifiable but not $(d-1)$-identifiable, and thus there must be some level set $\Gamma_r$ which is a flat of dimension $n-d-1$.
  \raft{What I mean is you can't describe $\Gamma_r$ using a flat of another dimension; revisit once we settle terminology}
